---
title: "GSS challange"
author: "Rabi Elkhoury"
date: "6/21/2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata) 
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

The question we are trying to answer with our data is:  
Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

## Now with The Most data (fewer variables gives more data points)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Based on the graphs alone, it seems that white citizens tend to favor the death penalty overwhelmingly (over 65%, according to the second graph).  This is to be contrasted to black citizens, who disapprove by a small margin, and other groups, which approve by a somewhat larger margin.  The data, thus far, has yielded mixed results.  

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Now with More data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```

Because the sample size of white persons surveyed is significantly higher compared to other groups, whites hold the highest "favor" and "oppose" proportions, according to the second table.  Examining the table corroborates thee findings from the earlier graphs; whites favor capital punishment among themselves.  Blacks seem to oppose it by slim margins, and "other" groups generally somewhat support it by larger percentages.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

## More data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```

```{r}
fisherXY(mydata$cappun,mydata$race)
```
 
Because the p-value with the initial Chi-Square test is so low, we will run a FisherExact test to gain more information. 


Running the FisherXY test yields p-values less than 0.05 between all groups (White/Black, WHITE/OTHER, BLACK/OTHER), implying that there is a significant difference between all of these groups.  The p-value indicates the probability of obtaining certain results wholly due to chance; the smaller it is, the less likely the data in question was "random".  With this statistic, we can reject the null hypothesis in favor of the alternate hypothesis.    
 
 
 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

When combining our assessments of our graphical, tabular, and inferential data, it is fair to assert that race affects one's opinion on capital punishment.  Clearly, white people seem to support capital punishment in proportions larger than black or other citizens (the percentages among the black and other groups are mixed, at best).  Much of this was established through the graphs and tables generated, but was solidified by the FisherXY test at the end.

Compared to branch2, the master branch omitted all tabulations of data that might include an "NA".  Because of this, there was more useful data that could be analyzed, thus leading to more accurate results.


